---
title: "EDS241: Assignment 2"
author: "Peter Menzies"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: no
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)


# load packages
librarian::shelf("stargazer", "here", "tidyr", "dplyr","stringr", "janitor", 
           "cowplot", "ggplot2", "tinytex", "datasets", "tibble", "tidyverse", 
            "estimatr", "car", "xtable", "readxl")

options(xtable.comment = FALSE)
# options(scipen=999) # not scientific notation
```

# Reading in data

```{r , include=TRUE}
df <- read_excel(here("data", "NBP.xls")) %>% 
  clean_names()
```

# (a) `dnox_masstons` distribution

```{r , include=TRUE}
dnox_dist <- ggplot(df, aes(x = dnox_masstons)) +
  geom_histogram()

dnox_dist
```


# (b) Creating an indicator (`D` = 1) representing `pct_black` above the sample median

```{r}
df_D <- df %>% 
  mutate("D" = case_when(
    pct_black > median(pct_black) ~ 1,
    pct_black <= median(pct_black) ~ 0
  ))

df_D1 <- df_D %>% 
  filter(D == 1)

avg_pct_D1 <- mean(df_D1$pct_black)
```

\noindent The average `pct_black` for counties above the median is `r avg_pct_D1`


# (c) Regression of `dnox_masstons` on `nbp`

```{r}
mdl_nbp <- lm_robust(dnox_masstons ~ nbp, df)
```

```{r, results='asis'}
mdl_nbp %>% 
  tidy %>% 
  xtable()
```

```{r, echo=FALSE, results='hide'}
mdl_nbp %>% summary()
```

```{r, echo=FALSE}
intercept <- mdl_nbp$coefficients[['(Intercept)']]
nbp <- mdl_nbp$coefficients[['nbp']]
```


\noindent The intercept represents the estimated decrease in tons of NOx between 2000 and 2008 in counties without `nbp` in effect.

\noindent The coefficient on `nbp` represents the estimated difference in change in tons of NOx between counties with `nbp` in effect and without `nbp` in effect between 2000 and 2008.

# (d) Linear regression of `dnox_masstons` on `nbp`, `D`, and `nbp` x `D`

```{r}
mdl_nbp_D <- lm_robust(dnox_masstons ~ nbp + D + nbp * D, df_D)
```

```{r, results='asis'}
mdl_nbp_D %>% 
  tidy %>% 
  xtable()
```

```{r, echo=FALSE, results='hide'}
mdl_nbp_D %>% summary()
```

```{r, echo=FALSE}
intercept <- mdl_nbp_D$coefficients[['(Intercept)']]
nbp <- mdl_nbp_D$coefficients[['nbp']]
D <- mdl_nbp_D$coefficients[['D']]
nbp_x_D <- mdl_nbp_D$coefficients[['nbp:D']]
```

\noindent The intercept tells us that counties without `nbp` in effect and with a `pct_black` less than or equal to the median saw an estimated decrease of `r abs(intercept)` tons of NOx between 2000 and 2008. 

\noindent The coefficient on `nbp` represents the estimated difference in change in tons of NOx in counties with `nbp` in effect and without `nbp` in effect that both have `pct_black` lower than the median.

\noindent The coefficient on `D` represents the estimated difference in change in tons of NOx in counties with `pct_black` higher than median and `pct_black` lower than median both without `nbp` in effect.

\noindent The coefficient on the interaction between `nbp` and `D` represents the estimated difference in change in tons of NOx associated with implementation of `nbp` in counties with `pct_black` higher than median and counties with `pct_black` lower than median. 



# (e) Predicted `dnox_masstons` in a county without NBP in effect and where `pct_black` is above the sample median

```{r}
x_vals <- tribble(~nbp, ~D,
                    0,   1)

pred_dnox <- predict(mdl_nbp_D,
                     newdata = x_vals,
                     se.fit = TRUE,
                     interval = 'confidence')
```

The predicted `dnox_masstons` in a county without NBP in effect and where `pct_black` is above the sample median is `r pred_dnox$fit[1]` tons of NOx. The 95% confidence interval for this prediction ranges from `r pred_dnox$fit[2]` to `r pred_dnox$fit[3]`.



