---
title: "EDS241: Assignment 1"
author: "Peter Menzies"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)


# load packages
librarian::shelf("stargazer", "here", "tidyr", "dplyr","stringr", "janitor", 
           "cowplot", "ggplot2", "tinytex", "datasets", "tibble", "tidyverse", 
           "patchwork", "estimatr", "car")

options(xtable.comment = FALSE)
# options(scipen=999) # not scientific notation
```

# Load data

\noindent Reading in the data and selecting the variables of interest

```{r , include=TRUE}
df <- read_csv(here("data", "CES4.csv")) %>% 
  clean_names() %>% 
  select(census_tract, total_population, california_county, low_birth_weight, pm2_5, poverty) %>% 
  rename(pm25 = pm2_5)
```

# (a) What is the average concentration of PM2.5 across all census tracts in California?

```{r , include=TRUE}
avg_pm25 <- mean(df$pm25)
```

\noindent The average concentration of PM2.5 across all census tracts in California is `r avg_pm25` $\mu g / m^3$

# (b) What county has the highest level of poverty in California?

```{r}
highest_poverty <- df %>% 
  filter(poverty == max(df$poverty, na.rm = TRUE))

highest_poverty_county <- highest_poverty$california_county
```

# (c) Make a histogram depicting the distribution of percent low birth weight and PM2.5.

```{r}
lbw_plot <- ggplot(df, aes(x = low_birth_weight)) +
  geom_histogram() +
  theme_minimal() +
  labs(x = "Percent low birth weight")

pm25_plot <- ggplot(df, aes(x = pm25)) +
  geom_histogram() +
  theme_minimal() +
  labs(x = "PM2.5")

lbw_plot + pm25_plot
```

# (d) OLS regression of `low_birth_weight` on `pm25`

```{r}
model_pm25 <- lm_robust(low_birth_weight ~ pm25, df)
```

```{r, results='asis', echo=FALSE}
model_pm25 %>%
  tidy() %>% 
  xtable::xtable(digits = 4, caption = "Linear regression of percent low birth weight on PM2.5")
```

```{r}
beta1_hat <- model_pm25$coefficients[["pm25"]]
```

\noindent The estimated slope coefficient is `r beta1_hat`. This tells us that an increase of 1 $\mu g / m^3$ of PM2.5 in a given California census tract is associated with an estimated increase of `r beta1_hat` in the percentage of births in that tract with weight less than 2500g. The effect of `pm25` on `low_birth_weight` *is* significant at the 5% level.

# (e) New air quality policy

```{r}
avg_new_lbw <- mean(df$low_birth_weight, na.rm = TRUE) - 2 * beta1_hat

lbw_ci_low <- mean(df$low_birth_weight, na.rm = TRUE) - 2 * model_pm25$conf.high[[2]]
lbw_ci_high <- mean(df$low_birth_weight, na.rm = TRUE) - 2 * model_pm25$conf.low[[2]]
```

\noindent The predicted average value of `low_birth_weight` given a reduction of 2 $\mu g / m^3$ of PM2.5 is `r avg_new_lbw`. The 95% confidence interval for this value is `r lbw_ci_low` to `r lbw_ci_high`, which means there is a 95% probability that this interval contains the true population estimate for `low_birth_weight` given a 2 $\mu g / m^3$ reduction in `pm25`

# (f)

```{r}
model_pm25_poverty <- lm_robust(low_birth_weight ~ pm25 + poverty, df)
```

```{r, results='asis', echo=FALSE}
model_pm25_poverty %>%
  tidy() %>% 
  xtable::xtable(digits = 4, caption = "Multiple linear regression of percent low birth weight on PM2.5 and poverty")
```

\noindent The estimated coefficient on `poverty` signifies that increasing the percentage of the population in the census tract living below twice the federal poverty line by 1% is associated with an estimated increase of `r model_pm25_poverty$coef[["poverty"]]` in `low_birth_weight`.

\noindent The estimated coefficient on `pm25` decreased from `r model_pm25$coefficients[["pm25"]]` in our original single variable model, to `r model_pm25_poverty$coefficients[["pm25"]]` in our multivariate model. This is because `pm25` and `poverty` are correlated (shown below)---and thus in our original model, the estimated coefficient on `pm25` was attempting to explain some of the variation in `low_birth_weight` that our second model shows to be associated with `poverty`.

```{r}
# showing correlation between pm25 and poverty
model_endogenous <- lm_robust(pm25 ~ poverty, df)
```

```{r, results='asis', echo=FALSE}
model_endogenous %>% 
  tidy() %>% 
  xtable::xtable(digits = 4, caption = "Linear regression of PM2.5 on poverty")
```

# (g) Linear hypothesis test

```{r}
lht <- linearHypothesis(model_pm25_poverty, "pm25=poverty", white.adjust = "hc2") 
```

```{r, results='asis', echo=FALSE}
lht %>% xtable::xtable(digits = 4,
                       caption = "Comparing effects of PM2.5 and poverty on percent low birth weight")
```

\noindent Based on the results of our linear hypothesis test, we can reject the null hypothesis that the effects of `pm25` and `poverty` on `low_birth_weight` are equal at the 0.1% level.
