---
title: "EDS241: Assignment 4"
author: "Peter Menzies"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: no
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)


# load packages
librarian::shelf("stargazer", "here", "tidyr", "dplyr","stringr", "janitor", 
           "cowplot", "ggplot2", "tinytex", "datasets", "tibble", "tidyverse", 
            "estimatr", "car", "xtable", "AER")

options(xtable.comment = FALSE)
```

# Reading in data

```{r , include=TRUE}
df <- read_csv(here("data", "EU_sardines.csv")) %>% 
  clean_names() %>% 
  mutate(log_price = log(price_euro_kg),
         log_volume = log(volume_sold_kg))
```

# (a) Bivariate regression of `log(volume_sold_kg)` on `log(price_euro_kg)`.

```{r}
mdl_a <- lm_robust(log_volume ~ log_price, df)
```

```{r, results='hide'}
mdl_a %>% summary()
elast_a <- mdl_a$coefficients["log_price"]
```

```{r, results='asis'}
mdl_a %>% 
  tidy() %>% 
  xtable()
```

```{r, results='hide'}
linearHypothesis(mdl_a, "log_price=-1", white.adjust = "hc2", test = "F")
```

```{r, results='asis'}
linearHypothesis(mdl_a, "log_price=-1", white.adjust = "hc2", test = "F") %>% 
  tidy %>% 
  xtable()
```


\noindent The estimated price elasticity of demand for sardines is `r elast_a`.

\noindent Based on our linear hypothesis test, we can reject the null hypothesis that the price elasticity of demand for sardines is equal to -1 with a p-value of less than 0.001.


# (b) Wind speed as an instrumental variable: first-stage regression of `log(price_euro_kg)` on `wind_m_s`

```{r}
mdl_b <- lm_robust(log_price ~ wind_m_s, df)
```

```{r, include=FALSE}
mdl_b %>% summary

coef_wind_b <- mdl_b$coefficients['wind_m_s']
fstat_b <- mdl_b$fstatistic[1]
```

```{r, results='asis'}
mdl_b %>% 
  tidy() %>% 
  xtable()
```




\noindent The estimated coefficient on `wind_m_s` represents the increase of `log(price_euro_kg)` given a 1 m/s increase in wind speed. The coefficient is positive---this makes sense as we would expect fish prices to increase with wind speeds due to higher wind speeds posing challenges for fishing vessels.

\noindent The first stage F-statistic, `r fstat_b`, is greater than 10 and thus `wind_m_s` is a relevant and non-weak instrument.


# (c) Two-stage least squares estimator of the price elasticity of demand for sardines using `wind_m_s` as an instrument for `log(price_euro_kg)`

```{r}
tsls_c <- ivreg(log_volume ~ log_price | wind_m_s, data = df)
```

```{r, results='hide'}
elast_c <- tsls_c$coefficients["log_price"]
tsls_c %>% summary()
```


```{r, results='asis'}
tsls_c %>% 
  tidy() %>% 
  xtable()
```



\noindent The estimated price elasticity of demand for sardines based on this TSLS estimator is `r elast_c`.



# (d) Two-stage least squares estimator of the price elasticity of demand for sardines using `wind_m_s` as an instrument for `log(price_euro_kg)` with fixed effects for `year`, `month`, and `country`.


```{r}
tsls_d <- ivreg(log_volume ~ log_price + as.factor(country) + as.factor(year) + as.factor(month) |
                  wind_m_s + as.factor(country) + as.factor(year) + as.factor(month), data = df)
```

```{r, results='hide', include=FALSE}
tsls_d %>% summary()
elast_d <- tsls_d$coefficients['log_price']
```

```{r, results='asis'}
tsls_d %>% 
  tidy %>% 
  xtable()
```


```{r}
mdl_d <- lm_robust(log_price ~ wind_m_s + 
                     as.factor(country) + 
                     as.factor(year) + 
                     as.factor(month), df)
```


```{r}
fstat_d <- mdl_d$fstatistic[1]
fstat_d
```


\noindent The estimated price elasticity based on this TSLS estimator is `r elast_d`. The F-statistic associated with wind as an instrumental variable with fixed effects for country, year, and month is `r fstat_d`, and thus the instruments are considered relevant and non-weak.






