---
title: "EDS241: Assignment 3"
author: "Peter Menzies"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: no
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)


# load packages
librarian::shelf("stargazer", "here", "tidyr", "dplyr","stringr", "janitor", 
           "cowplot", "ggplot2", "tinytex", "datasets", "tibble", "tidyverse", 
            "estimatr", "car", "xtable")

options(xtable.comment = FALSE)
```

# Reading in data

```{r , include=TRUE}
df <- read_csv(here("data", "SMOKING_EDS241.csv")) %>% 
  clean_names()
```

# (a) Unadjsted mean difference in birth weight of infants with smoking and nonsmoking mothers

```{r}
mean_wt_smoking <- mean(df$birthwgt[df$tobacco == 1])
mean_wt_nonsmoking <- mean(df$birthwgt[df$tobacco == 0])

wt_diff <- mean_wt_smoking - mean_wt_nonsmoking
```

```{r}
t.test(df$alcohol[df$tobacco == 1], df$alcohol[df$tobacco == 0])
```

\noindent The unadjusted mean difference in birth weight between infants with mothers who smoked during pregnancy and those with mothers who did not is `r wt_diff` grams.

\noindent This would correspond to the average treatment effect of maternal smoking during pregnancy on infant birth weight under the assumption that the treatment (mother smoking during pregnancy) is statistically independent of all relevant baseline characteristics. 

\noindent The mean difference in alcohol consumption during pregnancy between smoking and non-smoking mothers is statistically different from zero (shown above with a t-test), and thus we can't make the previous assumption regarding ATE.


# (b) Multiple linear regression of birth weight on smoking and all control variables

```{r}
mdl_b <- lm_robust(birthwgt ~ ., df)
```

```{r, results='asis'}
mdl_b %>% 
  tidy() %>% 
  xtable()
```


```{r, include=FALSE}
mdl_b

coef_b <- mdl_b$coefficients[["tobacco"]]
se_b <- mdl_b$std.error[["tobacco"]]
```



\noindent The estimated coefficient on tobacco in this model is `r coef_b` $\pm$ `r se_b` grams.


# (c) Estimating the effect of maternal smoking on birth weight using exact matching estimator

```{r}
df_c <- df %>% 
  mutate(mage = case_when(mage >= 34 ~ 1,
                          mage < 34 ~ 0)) %>% 
  mutate(meduc = case_when(meduc >= 16 ~ 1,
                           meduc < 16 ~ 0)) %>% 
  mutate(g = as.factor(paste0(mage, meduc, mblack, alcohol)))
```

```{r}
tia <- df_c %>%
  group_by(g, tobacco) %>% 
  summarise(n_obs = as.integer(n()),
            mean_wt = mean(birthwgt, na.rm = T)) %>%
  gather(variables, values, n_obs:mean_wt) %>% 
  mutate(variables = paste0(variables, "_", tobacco)) %>% 
  pivot_wider(id_cols = g, names_from = variables, values_from = values) %>% 
  ungroup() %>% 
  mutate(wt_diff = mean_wt_1 - mean_wt_0,
         w_ate = (n_obs_0 + n_obs_1) / (sum(n_obs_0) + sum(n_obs_1)),
         w_att = n_obs_1 / sum(n_obs_1)) %>% 
  mutate_if(is.numeric, round, 2)
```

```{r, results='asis'}
xtable(tia, digits = c(0, 0, 0, 0 , 2, 2, 2, 2, 2))
```

```{r}
ate = sum((tia$w_ate) * (tia$wt_diff))
ate

att = sum((tia$w_att) * (tia$wt_diff))
att
```


```{r}
lm_c <- lm_robust(birthwgt ~ tobacco + g, df_c)
```

```{r, results='asis'}
lm_c %>% 
  tidy() %>% 
  xtable()
```


```{r, include=FALSE}
lm_est_c <- lm_c$coefficients[["tobacco"]] 
```


\noindent The estimated average treatment effect of smoking during pregnancy on birth weight using the exact matching estimator is `r ate` grams compared with `r lm_est_c` grams for the analogous linear regression.


\newpage

\ 

# (d) Estimating the propensity score for maternal smoking using a logit estimator

```{r}
df_d <- df %>% 
  mutate(mage2 = mage ** 2)
```

```{r}
logit_mdl <- glm(tobacco ~ mage + mage2 + meduc + mblack + alcohol,
                 family = binomial(link = "logit"),
                 data = df_d)
```

```{r, results='asis'}
logit_mdl %>%
  tidy() %>% 
  xtable()
```

```{r}
eps <- predict(logit_mdl, type = "response")

ps_wgt <- (df_d$tobacco / eps) + ((1 - df_d$tobacco) / (1 - eps))
```


# (e) Using propensity score weighted regression to estimate the effect of maternal smoking on birth weight

```{r}
wls_mdl <- lm_robust(birthwgt ~ tobacco, df, weights = ps_wgt)
```

```{r, include=FALSE}
coef_wls <- wls_mdl$coefficients[['tobacco']]
```

```{r, results='asis'}
wls_mdl %>%
  tidy() %>% 
  xtable()
```

The estimated effect of maternal smoking on birth weight using propensity score weighted regression is `r coef_wls` grams.


