---
title: "EDS241: Assignment 3"
author: "Peter Menzies"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: no
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)


# load packages
librarian::shelf("stargazer", "here", "tidyr", "dplyr","stringr", "janitor", 
           "cowplot", "ggplot2", "tinytex", "datasets", "tibble", "tidyverse", 
            "estimatr", "car", "xtable", "readxl")

options(stargazer.header = FALSE)
```

# Reading in data

```{r , include=TRUE}
df <- read_csv(here("data", "SMOKING_EDS241.csv")) %>% 
  clean_names()
```

# (a) Mean difference in birth weight of infants with smoking and nonsmoking mothers

```{r , include=TRUE}
df_wt_diff <- df %>% 
  group_by(tobacco) %>% 
  summarize(mean_wt = mean(birthwgt))

wt_diff <- df_wt_diff$mean_wt[df_wt_diff$tobacco == 0] - df_wt_diff$mean_wt[df_wt_diff$tobacco == 1]
```

```{r}
t.test(df$alcohol[df$tobacco == 1], df$alcohol[df$tobacco == 0])
```

\noindent The unadjusted mean difference in birth weight of infants with smoking and nonsmoking mothers is `r wt_diff` grams.

\noindent This would correspond to the average treatment effect of maternal smoking during pregnancy on infant birth weight under the assumption that the treatment (mother smoking during pregnancy) is statistically independent of all control variables. 

\noindent The mean difference in alcohol consumption during pregnancy (one of the control variables) between smoking and non-smoking mothers is statistically different from zero (shown above with a t-test), and thus we can't make the previous assumption regarding ATE.


# (b) Multiple linear regression of birth weight on smoking and control variables

```{r}
mdl_b <- lm_robust(birthwgt ~ ., df)
```

```{r, results='asis'}
mdl_b %>% 
  starprep() %>% 
  stargazer(header = FALSE)
```


```{r, include=FALSE}
mdl_b

coef_b <- mdl_b$coefficients[["tobacco"]]
se_b <- mdl_b$std.error[["tobacco"]]
```



\noindent The estimated coefficient on tobacco in this model is `r coef_b` +/- `r se_b`


# (c) Estimating the effect of maternal smoking on birth weight using exact matching estimator

```{r}
df_c <- df %>% 
  mutate(mage = case_when(mage >= 34 ~ 1,
                          mage < 34 ~ 0)) %>% 
  mutate(meduc = case_when(meduc >= 16 ~ 1,
                           meduc < 16 ~ 0)) %>% 
  mutate(g = as.factor(paste0(mage, meduc, mblack, alcohol)))
```

```{r}
tia <- df_c %>%
  group_by(g, tobacco) %>% 
  summarise(n_obs = n(),
            mean_wt = mean(birthwgt, na.rm = T)) %>%
  gather(variables, values, n_obs:mean_wt) %>% 
  mutate(variables = paste0(variables, "_", tobacco)) %>% 
  pivot_wider(id_cols = g, names_from = variables, values_from = values) %>% 
  ungroup() %>% 
  mutate(wt_diff = mean_wt_1 - mean_wt_0,
         w_ate = (n_obs_0 + n_obs_1) / (sum(n_obs_0) + sum(n_obs_1)),
         w_att = n_obs_1 / sum(n_obs_1)) %>% 
  mutate_if(is.numeric, round, 2)
```

```{r, results='asis'}
stargazer(tia, summary = FALSE, digits = 2, header = FALSE)
```

```{r}
ate = sum((tia$w_ate) * (tia$wt_diff))
ate

att = sum((tia$w_att) * (tia$wt_diff))
att
```


```{r}
lm_c <- lm_robust(birthwgt ~ tobacco + g, df_c)
```

```{r, results='hide'}
lm_c$coefficients %>%
  stargazer(header = FALSE, digits = 2, column.sep.width = "0.5pt")
```

\begin{table}[!htbp] \centering 
  \caption{} 
  \label{} 
\begin{tabular}{@{\extracolsep{0.5pt}} ccccccccccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
(Intercept) & tobacco & g0001 & g0010 & g0011 & g0100 & g0101 & g0110 & g0111 \\ 
\hline \\[-1.8ex] 
$3,445.87$ & $$-$226.25$ & $$-$63.12$ & $$-$241.84$ & $$-$384.01$ & $37.81$ & $88.51$ & $$-$120.78$ & $$-$219.20$ \\ 
\\
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
g1000 & g1001 & g1010 & g1011 & g1100 & g1101 & g1110 & g1111 \\
\hline \\[-1.8ex] 
$10.36$ & $$-$102.85$ & $$-$251.69$ & $$-$443.86$ & $40.82$ & $26.74$ & $$-$146.19$ & $$-$185.75$$$ \\
\end{tabular} 
\end{table} 
\ 


```{r, include=FALSE}
lm_est_c <- lm_c$coefficients[["tobacco"]] 
```


\noindent The estimated average treatment effect of smoking on birth weight using the exact matching estimator is `r ate` compared with `r lm_est_c` for the analagous linear regression.


# (d) Estimating the propensity score for maternal smoking using a logit estimator

```{r}
df_d <- df %>% 
  mutate(mage2 = mage ^ 2,
         wt_scaled = scale(birthwgt))
```

```{r}
logit_mdl <- glm(tobacco ~ mage + mage2 + meduc + mblack + alcohol, family = binomial(link = "logit"), df_d)
```

```{r, results='asis'}
logit_mdl %>% stargazer(header = FALSE)
```

```{r}
eps <- predict(logit_mdl, type = "response")

ps_wgt <- (df_d$tobacco / eps) + ((1 - df_d$tobacco) / (1 - eps))
```


# (e) Using the propensity score weighted regression to estimate the effect of maternal smoking on birth weight

```{r}
wls_mdl <- lm_robust(birthwgt ~ tobacco, df, weights = ps_wgt)
```

```{r, results='asis'}
wls_mdl %>%
  starprep() %>% 
  stargazer(header = FALSE)
```




